---
title: "1-1 R-R"
output: html_document
date: "2025-07-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(mgcv)
library(tidyr)
library(ggforce)
library(GGally)
library(minpack.lm)
```


#______________________________check if there are outliers_____________________________________
#### 159 -> 252
# extract experimental data
```{r}
datafolder = "C:/Users/ZTian/Desktop/RP1/data/moduleTable.txt"
module_table <- read.delim(datafolder)
module_table <- module_table %>%
  mutate(conc_level = case_when(
    conc_level == 1 ~ 250,
    conc_level == 2 ~ 500,
    conc_level == 3 ~ 1000
  )) %>%
    filter(!(module == "hRPTECTERT1_97" & conc_level == 1000 &
           startsWith(as.character(eg_score), "-0.527544"))) %>%
    filter(!(module == "hRPTECTERT1_264" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.035441"))) %>%
    filter(!(module == "hRPTECTERT1_71" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.647139")))
KE1 <- module_table %>% dplyr::filter(
    module == 'hRPTECTERT1_159' & # Increase, Cytotoxicity
      time == 24 &
      experiment %in% c(
        'uploadFile1_GENTAMYCINluhumanrptectert1',
        'uploadFile2_GENTAMYCINluhumanrptectert1',
        'uploadFile3_GENTAMYCINluhumanrptectert1'
      )
  )
  print(KE1)

  KE2 <- module_table %>% dplyr::filter(
    module == 'hRPTECTERT1_252' & # Increase, Regenerative cell proliferation
      time == 24 &
      experiment %in% c(
        'uploadFile1_GENTAMYCINluhumanrptectert1',
        'uploadFile2_GENTAMYCINluhumanrptectert1',
        'uploadFile3_GENTAMYCINluhumanrptectert1'
      )
  )
  print(KE2)
  
  # simplify these data frame and merge them
  KE1_filtered <- KE1 %>%
    select(experiment, conc_level, eg_score) %>%
    rename(eg_score1 = eg_score)
  
  KE2_filtered <- KE2 %>%
    select(experiment, conc_level, eg_score) %>%
    rename(eg_score2 = eg_score)
  
  merged_data <- KE1_filtered %>%
  full_join(KE2_filtered, by = c("experiment", "conc_level")) 
  print(merged_data)
```

# PLOT KE1 VS KE2 USING EXPERIMENTAL DATA
```{r}
summarized_moduletable = module_table %>% filter(time == 24) %>%
  group_by(module, conc_level) %>%
  summarize(mean_eg = mean(eg_score, na.rm = TRUE), SD_eg = sd(eg_score, na.rm = TRUE)) %>%
  pivot_wider(names_from = module, values_from = c(mean_eg, SD_eg))

p <- ggplot() +
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score1, y = eg_score2, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_159, y = mean_eg_hRPTECTERT1_252, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_159,
                    ymin = mean_eg_hRPTECTERT1_252 - SD_eg_hRPTECTERT1_252,
                    ymax = mean_eg_hRPTECTERT1_252 + SD_eg_hRPTECTERT1_252,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_252,
                     xmin = mean_eg_hRPTECTERT1_159 - SD_eg_hRPTECTERT1_159,
                     xmax = mean_eg_hRPTECTERT1_159 + SD_eg_hRPTECTERT1_159,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Increase, Cytotoxicity VS Increase, Regenerative cell proliferation", x = "Increase, Cytotoxicity", y = "Increase, Regenerative cell proliferation", color = "DOSE") +
  theme_minimal()
print(p)
```

#### 277 -> 229
# extract experimental data
```{r}
datafolder = "C:/Users/ZTian/Desktop/RP1/data/moduleTable.txt"
module_table <- read.delim(datafolder)
module_table <- module_table %>%
  mutate(conc_level = case_when(
    conc_level == 1 ~ 250,
    conc_level == 2 ~ 500,
    conc_level == 3 ~ 1000
  )) %>%
    filter(!(module == "hRPTECTERT1_97" & conc_level == 1000 &
           startsWith(as.character(eg_score), "-0.527544"))) %>%
    filter(!(module == "hRPTECTERT1_264" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.035441"))) %>%
    filter(!(module == "hRPTECTERT1_71" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.647139")))
KE1 <- module_table %>% dplyr::filter(
    module == 'hRPTECTERT1_277' & # Increase, Cytotoxicity
      time == 24 &
      experiment %in% c(
        'uploadFile1_GENTAMYCINluhumanrptectert1',
        'uploadFile2_GENTAMYCINluhumanrptectert1',
        'uploadFile3_GENTAMYCINluhumanrptectert1'
      )
  )
  print(KE1)

  KE2 <- module_table %>% dplyr::filter(
    module == 'hRPTECTERT1_229' & # Increase, Regenerative cell proliferation
      time == 24 &
      experiment %in% c(
        'uploadFile1_GENTAMYCINluhumanrptectert1',
        'uploadFile2_GENTAMYCINluhumanrptectert1',
        'uploadFile3_GENTAMYCINluhumanrptectert1'
      )
  )
  print(KE2)
  
  # simplify these data frame and merge them
  KE1_filtered <- KE1 %>%
    select(experiment, conc_level, eg_score) %>%
    rename(eg_score1 = eg_score)
  
  KE2_filtered <- KE2 %>%
    select(experiment, conc_level, eg_score) %>%
    rename(eg_score2 = eg_score)
  
  merged_data <- KE1_filtered %>%
  full_join(KE2_filtered, by = c("experiment", "conc_level")) 
  print(merged_data)
```
# PLOT KE1 VS KE2 USING EXPERIMENTAL DATA
```{r}
summarized_moduletable = module_table %>% filter(time == 24) %>%
  group_by(module, conc_level) %>%
  summarize(mean_eg = mean(eg_score, na.rm = TRUE), SD_eg = sd(eg_score, na.rm = TRUE)) %>%
  pivot_wider(names_from = module, values_from = c(mean_eg, SD_eg))

p <- ggplot() +
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score1, y = eg_score2, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_277, y = mean_eg_hRPTECTERT1_229, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_277,
                    ymin = mean_eg_hRPTECTERT1_229 - SD_eg_hRPTECTERT1_229,
                    ymax = mean_eg_hRPTECTERT1_229 + SD_eg_hRPTECTERT1_229,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_229,
                     xmin = mean_eg_hRPTECTERT1_277 - SD_eg_hRPTECTERT1_277,
                     xmax = mean_eg_hRPTECTERT1_277 + SD_eg_hRPTECTERT1_277,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Inhibition, mtETC Chain Complexes VS Increase, NF-kb", x = "Inhibition, mtETC Chain Complexes", y = "Increase, NF-kb", color = "DOSE") +
  theme_minimal()
print(p)
```

#### 50 -> 264
# extract experimental data
```{r}
datafolder = "C:/Users/ZTian/Desktop/RP1/data/moduleTable.txt"
module_table <- read.delim(datafolder)
module_table <- module_table %>%
  mutate(conc_level = case_when(
    conc_level == 1 ~ 250,
    conc_level == 2 ~ 500,
    conc_level == 3 ~ 1000
  )) %>%
    filter(!(module == "hRPTECTERT1_97" & conc_level == 1000 &
           startsWith(as.character(eg_score), "-0.527544"))) %>%
    filter(!(module == "hRPTECTERT1_264" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.035441"))) %>%
    filter(!(module == "hRPTECTERT1_71" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.647139")))
KE1 <- module_table %>% dplyr::filter(
    module == 'hRPTECTERT1_50' & # Increase, Cytotoxicity
      time == 24 &
      experiment %in% c(
        'uploadFile1_GENTAMYCINluhumanrptectert1',
        'uploadFile2_GENTAMYCINluhumanrptectert1',
        'uploadFile3_GENTAMYCINluhumanrptectert1'
      )
  )
  print(KE1)

  KE2 <- module_table %>% dplyr::filter(
    module == 'hRPTECTERT1_264' & # Increase, Regenerative cell proliferation
      time == 24 &
      experiment %in% c(
        'uploadFile1_GENTAMYCINluhumanrptectert1',
        'uploadFile2_GENTAMYCINluhumanrptectert1',
        'uploadFile3_GENTAMYCINluhumanrptectert1'
      )
  )
  print(KE2)
  
  # simplify these data frame and merge them
  KE1_filtered <- KE1 %>%
    select(experiment, conc_level, eg_score) %>%
    rename(eg_score1 = eg_score)
  
  KE2_filtered <- KE2 %>%
    select(experiment, conc_level, eg_score) %>%
    rename(eg_score2 = eg_score)
  
  merged_data <- KE1_filtered %>%
  full_join(KE2_filtered, by = c("experiment", "conc_level")) 
  print(merged_data)
```

# PLOT KE1 VS KE2 USING EXPERIMENTAL DATA
```{r}
summarized_moduletable = module_table %>% filter(time == 24) %>%
  group_by(module, conc_level) %>%
  summarize(mean_eg = mean(eg_score, na.rm = TRUE), SD_eg = sd(eg_score, na.rm = TRUE)) %>%
  pivot_wider(names_from = module, values_from = c(mean_eg, SD_eg))

p <- ggplot() +
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score1, y = eg_score2, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_50, y = mean_eg_hRPTECTERT1_264, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_50,
                    ymin = mean_eg_hRPTECTERT1_264 - SD_eg_hRPTECTERT1_264,
                    ymax = mean_eg_hRPTECTERT1_264 + SD_eg_hRPTECTERT1_264,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_264,
                     xmin = mean_eg_hRPTECTERT1_50 - SD_eg_hRPTECTERT1_50,
                     xmax = mean_eg_hRPTECTERT1_50 + SD_eg_hRPTECTERT1_50,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Increased, oxidative stress VS Increased, Lipid peroxidation", x = "Increased, oxidative stress", y = "Increased, Lipid peroxidation", color = "DOSE") +
  theme_minimal()
print(p)
```

#__________________________________________model________________________________________________
##### 159 -> 252
##KE1: Increase, Cytotoxicity, 159
##KE2: Increase, Regenerative cell proliferation, 252
```{r}
# define MM.2 manually
MM2_function <- function(dose, d, e) {
  d/(1 + (e/dose))
}
# define MM.3 manually
MM3_function <- function(dose, c, d, e) {
  c + (d - c)/(1 + (e/dose))
}
# define the range of dose
# # change to smaller range when plotting!!!! otherwise it's too large, the plots won't be beautifal
doses <- seq(100, 5000, by = 50)
# doses <- 10^seq(log10(100), log10(500000), length.out = 200)
# ceate an empty data frame
response_df <- data.frame(dose = doses)

# stimulate KE1 data(159), using MM.2 model
response_df$KE1 <- MM2_function(doses, d = 7.46, e = 2206.22)

# stimulate KE2 data(252), using MM.3 model
response_df$KE2 <- MM3_function(doses, c = -0.78, d = 25.68, e = 8775.51)

rownames(response_df) <- response_df[, 1] # set the first column as row names(dose)
response_df <- response_df[, -1] # delete the first column

print(response_df)
```

```{r}
pairs(response_df, pch = 18, col = "steelblue")
ggpairs(response_df)
```

# fit the model
```{r}
# model <- nlsLM(KE2 ~ a * exp(b * KE1) - c, data = response_df, start = list(a = 2.5, b = 0.3, c = 3), control = nls.control(maxiter = 800, warnOnly = TRUE))
# summary(model)
model <- nlsLM(KE2 ~ a * (KE1/(b + KE1)) - c, data = response_df, start = list(a = -1, b = 5, c = 0.4), control = nls.control(maxiter = 500, warnOnly = TRUE))
summary(model)
```
#### The following two chunks were used to check how well the fitted model matches the data used for fitting.
#### when dealing with complex models that are hard to fit, these plots help in selecting initial parameter estimates that closely align with the data used for fitting.
```{r}
response_function <- function(KE1, a, b, c){
  # a * exp(b * KE1)
  # a *(KE1/(b + KE1))
  # a * exp(b * KE1) - c
  a * (KE1/(b + KE1)) - c
}

# set parameters
a <- -8.886e+00
b <- -9.965e+00
c <- 7.800e-01

# predict KE2
response_df <- response_df %>%
  mutate(p_KE2 = response_function(KE1, a, b, c))

print(response_df)
```
```{r}
## determine model and estimate initial parameters
# plot: KE1 vs KE2
p <- ggplot() +
  # predicted data
  geom_point(data = response_df, aes(x = KE1, y = KE2, color = "drm", shape = "Predicted"), size = 2) +
  geom_smooth(data = response_df, aes(x = KE1, y = p_KE2, colour = "rrm", linetype = "rrm"), method = "gam", se = FALSE, color = "blue") + # smoother method
  # experimental data
  geom_point(data = response_df, aes(x = KE1, y = KE2, color = "Predicted from drm", shape = "Predicted from drm"), size = 2) +
  scale_color_manual(name = "Data Type", values = c("Predicted" = "red", "Predicted from drm" = "red")) +
  scale_shape_manual(name = "Data Type", values = c("Predicted" = 16, "Predicted from drm" = 1)) +
  theme_minimal() +
  #facet_zoom(y = p_KE3 < 30) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  labs(title = "KE1 vs KE2", x = "KE1", y = "KE2")
print(p)
```

```{r}
## 159 - 252
##KE1: Increase, Cytotoxicity, 159
##KE2: Increase, Regenerative cell proliferation, 252
## final plot
summarized_moduletable = module_table %>% filter(time == 24) %>%
  # # remove outlier from error bar
  filter(!(module == "hRPTECTERT1_97" & conc_level == 1000 &
           startsWith(as.character(eg_score), "-0.527544"))) %>%
  filter(!(module == "hRPTECTERT1_264" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.035441"))) %>%
  filter(!(module == "hRPTECTERT1_71" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.647139"))) %>%
  group_by(module, conc_level) %>%
  summarize(mean_eg = mean(eg_score, na.rm = TRUE), SD_eg = sd(eg_score, na.rm = TRUE)) %>%
  pivot_wider(names_from = module, values_from = c(mean_eg, SD_eg))
# # remove outlier from experimental data points
#   merged_data <- merged_data[-7, ]
#   # filter(!(eg_score2 == -0.5275449 & eg_score3 == 0.6471399)), cannot delete the row due to floating-point precision

# plot1: KE1 vs KE3
p1 <- ggplot() +
  # # predicted data
  geom_point(data = response_df, aes(x = KE1, y = KE2, color = "drm"), size = 2) +
  geom_smooth(data = response_df, aes(x = KE1, y = p_KE2, colour = "rrm", linetype = "rrm"), method = "gam", se = FALSE, color = "blue") +
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score1, y = eg_score2, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_159, y = mean_eg_hRPTECTERT1_252, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_159,
                    ymin = mean_eg_hRPTECTERT1_252 - SD_eg_hRPTECTERT1_252,
                    ymax = mean_eg_hRPTECTERT1_252 + SD_eg_hRPTECTERT1_252,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_252,
                     xmin = mean_eg_hRPTECTERT1_159 - SD_eg_hRPTECTERT1_159,
                     xmax = mean_eg_hRPTECTERT1_159 + SD_eg_hRPTECTERT1_159,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Increase, Cytotoxicity VS Increase, Regenerative cell proliferation", x = "Increase, Cytotoxicity", y = "Increase, Regenerative cell proliferation", color = "DOSE") +
  # facet_zoom(y = p_KE3 < 25) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  theme_minimal()
print(p1)
```
### save
```{r}
ggsave(filename = "../data/R_R/1_1 R_R/p1.pdf",
       plot = p1,
       width = 8,
       height = 6,
       dpi = 300)
```



#### 277 -> 229
##KE1: Inhibition, mtETC Chain Complexes, 277
##KE2: Increase, NF-kb, 252
```{r}
# define MM.2 manually
MM2_function <- function(dose, d, e) {
  d/(1 + (e/dose))
}
# define MM.3 manually
MM3_function <- function(dose, c, d, e) {
  c + (d - c)/(1 + (e/dose))
}
# define LL.3 manually
LL3_function <- function(dose, b, d, e) {
  d / (1 + exp(b * (log(dose) - log(e))))
}
# define the range of dose
# change to smaller range when plotting!!!! otherwise it's too large, the plots won't be beautifal
doses <- seq(100, 5000, by = 50)
# doses <- 10^seq(log10(100), log10(500000), length.out = 400)
# ceate an empty data frame
response_df <- data.frame(dose = doses)

# stimulate KE1 data(277), using MM.2 model
response_df$KE1 <- MM2_function(doses, d = -33.31, e = 14683.08)

# stimulate KE2 data(229), using LL.3 model
response_df$KE2 <- LL3_function(doses, b = -6.94, d = 3.27, e = 426.37)

rownames(response_df) <- response_df[, 1] # set the first column as row names(dose)
response_df <- response_df[, -1] # delete the first column

print(response_df)
```

```{r}
pairs(response_df, pch = 18, col = "steelblue")
ggpairs(response_df)
```

# fit the model
```{r}
model <- nlsLM(KE2 ~ a / (1 + (b / KE1)^n), data = response_df, start = list(a = 3.2, b = -2, n = 7.63), control = nls.control(maxiter = 500, warnOnly = TRUE))
summary(model)
```

```{r}
response_function <- function(KE1, a, b, n){
  a / (1 + (b / KE1)^n)
}
# set parameters
a <- 3.27
b <- -0.94
n <- 7.14

# predict KE2
response_df <- response_df %>%
  mutate(p_KE2 = response_function(KE1, a, b, n))

print(response_df)
```
# final plotting
```{r}
#### 277 -> 229
##KE1: Inhibition, mtETC Chain Complexes, 277
##KE2: Increase, NF-kb, 229
## final plot
summarized_moduletable = module_table %>% filter(time == 24) %>%
  # # remove outlier from error bar
  filter(!(module == "hRPTECTERT1_97" & conc_level == 1000 &
           startsWith(as.character(eg_score), "-0.527544"))) %>%
  filter(!(module == "hRPTECTERT1_264" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.035441"))) %>%
  filter(!(module == "hRPTECTERT1_71" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.647139"))) %>%
  group_by(module, conc_level) %>%
  summarize(mean_eg = mean(eg_score, na.rm = TRUE), SD_eg = sd(eg_score, na.rm = TRUE)) %>%
  pivot_wider(names_from = module, values_from = c(mean_eg, SD_eg))
# # remove outlier from experimental data points
#   merged_data <- merged_data[-7, ]
#   # filter(!(eg_score2 == -0.5275449 & eg_score3 == 0.6471399)), cannot delete the row due to floating-point precision

# plot1: KE1 vs KE3
p2 <- ggplot() +
  # # predicted data
  geom_point(data = response_df, aes(x = KE1, y = KE2, color = "drm"), size = 2) +
  geom_smooth(data = response_df, aes(x = KE1, y = p_KE2, colour = "rrm", linetype = "rrm"), method = "gam", formula = y ~ s(x, k =70), se = FALSE, color = "blue") +
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score1, y = eg_score2, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_277, y = mean_eg_hRPTECTERT1_229, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_277,
                    ymin = mean_eg_hRPTECTERT1_229 - SD_eg_hRPTECTERT1_229,
                    ymax = mean_eg_hRPTECTERT1_229 + SD_eg_hRPTECTERT1_229,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_229,
                     xmin = mean_eg_hRPTECTERT1_277 - SD_eg_hRPTECTERT1_277,
                     xmax = mean_eg_hRPTECTERT1_277 + SD_eg_hRPTECTERT1_277,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Inhibition, mtETC Chain Complexes VS Increase, NF-kb", x = "Inhibition, mtETC Chain Complexes", y = "Increase, NF-kb", color = "DOSE") +
  # facet_zoom(y = p_KE3 < 25) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  theme_minimal()
print(p2)
```
### save
```{r}
ggsave(filename = "../data/R_R/1_1 R_R/p2.pdf",
       plot = p2,
       width = 8,
       height = 6,
       dpi = 300)
```




#### 50 -> 264
## KE1: Increased, oxidative stress, 50
## KE2: Increased, Lipid peroxidation, 264
```{r}
# define MM.2 manually
MM2_function <- function(dose, d, e) {
  d/(1 + (e/dose))
}
# define MM.3 manually
MM3_function <- function(dose, c, d, e) {
  c + (d - c)/(1 + (e/dose))
}
# define the range of dose
# change to smaller range when plotting!!!! otherwise it's too large, the plots won't be beautifal
doses <- seq(100, 5000, by = 50)
# doses <- 10^seq(log10(100), log10(5000000), length.out = 400)
# ceate an empty data frame
response_df <- data.frame(dose = doses)

# stimulate KE1 data(50), using MM.2 model
response_df$KE1 <- MM2_function(doses, d = 3.32, e = 437.75)

# stimulate KE2 data(264), using MM.3 model
response_df$KE2 <- MM3_function(doses, c = -0.58, d = 200.7, e = 89671.97)

rownames(response_df) <- response_df[, 1] # set the first column as row names(dose)
response_df <- response_df[, -1] # delete the first column

print(response_df)
```

```{r}
pairs(response_df, pch = 18, col = "steelblue")
ggpairs(response_df)
```

# fit the model
```{r}
model <- nlsLM(KE2 ~ a * (KE1/(b + KE1)) - c, data = response_df, start = list(a = -1, b = 5, c = 0.4), control = nls.control(maxiter = 500, warnOnly = TRUE))
summary(model)
```

```{r}
response_function <- function(KE1, a, b, c){
  a * (KE1/(b + KE1)) - c
}
# set parameters
a <- -9.874e-01
b <- -3.336e+00
c <- 5.800e-01

# predict KE2
response_df <- response_df %>%
  mutate(p_KE2 = response_function(KE1, a, b, c))

print(response_df)
```
```{r}
## determine model and estimate initial parameters
# plot: KE1 vs KE2
p <- ggplot() +
  # predicted data
  geom_point(data = response_df, aes(x = KE1, y = p_KE2, color = "Predicted", shape = "Predicted"), size = 2) +
  geom_smooth(data = response_df, aes(x = KE1, y = p_KE2), method = "gam", se = FALSE, color = "blue") + # smoother method
  # experimental data
  geom_point(data = response_df, aes(x = KE1, y = KE2, color = "Predicted from drm", shape = "Predicted from drm"), size = 2) +
  scale_color_manual(name = "Data Type", values = c("Predicted" = "red", "Predicted from drm" = "red")) +
  scale_shape_manual(name = "Data Type", values = c("Predicted" = 16, "Predicted from drm" = 1)) +
  theme_minimal() +
  #facet_zoom(y = p_KE3 < 30) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  labs(title = "KE1 vs KE2", x = "KE1", y = "KE2")
print(p)
```

# final plotting
```{r}
#### 50 -> 264
## KE1: Increased, oxidative stress, 50
## KE2: Increased, Lipid peroxidation, 264
## final plot
summarized_moduletable = module_table %>% filter(time == 24) %>%
  # # remove outlier from error bar
  filter(!(module == "hRPTECTERT1_97" & conc_level == 1000 &
           startsWith(as.character(eg_score), "-0.527544"))) %>%
  filter(!(module == "hRPTECTERT1_264" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.035441"))) %>%
  filter(!(module == "hRPTECTERT1_71" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.647139"))) %>%
  group_by(module, conc_level) %>%
  summarize(mean_eg = mean(eg_score, na.rm = TRUE), SD_eg = sd(eg_score, na.rm = TRUE)) %>%
  pivot_wider(names_from = module, values_from = c(mean_eg, SD_eg))
# # remove outlier from experimental data points
#   merged_data <- merged_data[-7, ]
#   # filter(!(eg_score2 == -0.5275449 & eg_score3 == 0.6471399)), cannot delete the row due to floating-point precision

# plot1: KE1 vs KE3
p3 <- ggplot() +
  # # predicted data
  geom_point(data = response_df, aes(x = KE1, y = KE2, color = "drm"), size = 2) +
  geom_smooth(data = response_df, aes(x = KE1, y = p_KE2, colour = "rrm", linetype = "rrm"), method = "gam", formula = y ~ s(x, k =70), se = FALSE, color = "blue") +
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score1, y = eg_score2, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_50, y = mean_eg_hRPTECTERT1_264, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_50,
                    ymin = mean_eg_hRPTECTERT1_264 - SD_eg_hRPTECTERT1_264,
                    ymax = mean_eg_hRPTECTERT1_264 + SD_eg_hRPTECTERT1_264,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_264,
                     xmin = mean_eg_hRPTECTERT1_50 - SD_eg_hRPTECTERT1_50,
                     xmax = mean_eg_hRPTECTERT1_50 + SD_eg_hRPTECTERT1_50,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Increased, oxidative stress VS Increased, Lipid peroxidation", x = "Increased, oxidative stress", y = "Increased, Lipid peroxidation", color = "DOSE") +
  # facet_zoom(y = p_KE3 < 25) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  theme_minimal()
print(p3)
```

### save
```{r}
ggsave(filename = "../data/R_R/1_1 R_R/p3.pdf",
       plot = p3,
       width = 8,
       height = 6,
       dpi = 300)
```