---
title: "R-R: 277 + 239 -> 50"
output: html_document
date: "2025-07-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(ggplot2)
library(dplyr)
```
### module 277 + module 239 = module 50
# KE1: Inhibition, mtETC Chain Complexes--277
# KE2: Binding, Thiol/seleno proteins--239
# KE3: Increased, oxidative stress--50
### use data predicted from dose-response models
```{r}
# 277, 50, 239: MM.2
# define MM.2 manually
MM2_function <- function(dose, d, e) {
  d/(1 + (e/dose))
}
# define the range of dose
# doses <- seq(100, 1000000, by = 50)
# doses <- 10^seq(log10(100), log10(1000000), length.out = 500)
# change to smaller range when plotting!!!! otherwise it's too large, the plots won't be beautifal
doses <- 10^seq(log10(100), log10(5000), length.out = 50)

# ceate an empty data frame
response_df <- data.frame(dose = doses)

# stimulate KE1 data(277), using MM.2 model
response_df$KE1 <- MM2_function(doses, d = -33.31, e = 14683.08)

# stimulate KE2 data(239), using MM.2 model
response_df$KE2 <- MM2_function(doses, d = -1.79, e = 194.3)

# stimulate KE3 data(50), using MM.2 model
response_df$KE3 <- MM2_function(doses, d = 3.32, e = 437.75)

rownames(response_df) <- response_df[, 1] # set the first column as row names(dose)
response_df <- response_df[, -1] # delete the first column

print(response_df)
```

# examine the data, check if the predictor variables have a linear association with the response variable
```{r}
pairs(response_df, pch = 18, col = "steelblue")
ggpairs(response_df)
```

# fit the model
```{r}
model <- nls(KE3 ~ a * (k / (k + KE1)) + c * (d / (d + KE2)), data = response_df, start = list(a = 1, k = -1.2, c = 1, d = -1), control = nls.control(maxiter = 500, warnOnly = TRUE))
summary(model)
```


#### The following two chunks were used to check how well the fitted model matches the data used for fitting.
#### when dealing with complex models that are hard to fit, these plots help in selecting initial parameter estimates that closely align with the data used for fitting.
```{r}
# define function
KE3_function <- function(KE1, KE2, a, k, c, d) {
  a * (k / (k + KE1)) + c * (d / (d + KE2))
}

# set parameters
a <- -3.422e+00
k <- -1.024e+00
c <- 3.422e+00
d <- -1.364e+07

# predict KE3
response_df <- response_df %>%
  mutate(p_KE3 = KE3_function(KE1, KE2, a, k, c, d))

print(response_df)
```
```{r}
# plot1: KE1 vs KE3
p1 <- ggplot() +
  # predicted data
  geom_point(data = response_df, aes(x = KE1, y = p_KE3, color = "Predicted", shape = "Predicted"), size = 2) +
  # geom_smooth(data = response_df, aes(x = KE1, y = p_KE3), method = "loess", se = FALSE, color = "blue") +
  geom_smooth(data = response_df, aes(x = KE1, y = p_KE3), method = "gam", se = FALSE, color = "blue") + # smoother method
  # experimental data
  geom_point(data = response_df, aes(x = KE1, y = KE3, color = "Predicted from drm", shape = "Predicted from drm"), size = 2) +
  scale_color_manual(name = "Data Type", values = c("Predicted" = "red", "Predicted from drm" = "red")) +
  scale_shape_manual(name = "Data Type", values = c("Predicted" = 16, "Predicted from drm" = 1)) +
  theme_minimal() +
  # facet_zoom(y = p_KE3 < 10) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  labs(title = "KE1 vs KE3", x = "KE1", y = "KE3")

# plot2: KE2 vs KE3
p2 <- ggplot(response_df, aes(x = KE2)) +
  # predicted data
   geom_point(data = response_df, aes(x = KE2, y = p_KE3, color = "Predicted", shape = "Predicted"), size = 2) +
  # geom_smooth(data = response_df, aes(x = KE2, y = p_KE3), method = "loess", se = FALSE, color = "blue") +
  geom_smooth(data = response_df, aes(x = KE2, y = p_KE3), method = "gam", se = FALSE, color = "blue") +
  # experimental data
  geom_point(data = response_df, aes(x = KE2, y = KE3, color = "Predicted from drm", shape = "Predicted from drm"), size = 2) +
  scale_color_manual(name = "Data Type", values = c("Predicted" = "red", "Predicted from drm" = "red")) +
  scale_shape_manual(name = "Data Type", values = c("Predicted" = 16, "Predicted from drm" = 1)) +
  theme_minimal() +
  # facet_zoom(y = p_KE3 < 10) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  labs(title = "KE2 vs KE3", x = "KE2", y = "KE3")

# print plots
print(p1)
print(p2)
```

### final plotting
```{r}
### module 277 + module 239 = module 50
# KE1: Inhibition, mtETC Chain Complexes--277
# KE2: Binding, Thiol/seleno proteins--239
# KE3: Increased, oxidative stress--50
# prepare the data frame used to draw error bars
# RStudio's table view (View()) only displays rounded numbers, not the actual stored values.
summarized_moduletable = module_table %>% filter(time == 24) %>%
  # # remove outlier from error bar
  # filter(!(module == "hRPTECTERT1_97" & conc_level == 1000 & 
  #          startsWith(as.character(eg_score), "-0.527544"))) %>%
  # filter(!(module == "hRPTECTERT1_264" & conc_level == 1000 & 
  #          startsWith(as.character(eg_score), "0.035441"))) %>%
  # filter(!(module == "hRPTECTERT1_71" & conc_level == 1000 & 
  #          startsWith(as.character(eg_score), "0.647139"))) %>%
  group_by(module, conc_level) %>%
  summarize(mean_eg = mean(eg_score, na.rm = TRUE), SD_eg = sd(eg_score, na.rm = TRUE)) %>%
  pivot_wider(names_from = module, values_from = c(mean_eg, SD_eg))
# # remove outlier from experimental data points
#   merged_data <- merged_data[-7, ]
#   # filter(!(eg_score2 == -0.5275449 & eg_score3 == 0.6471399)), cannot delete the row due to floating-point precision

# plot1: KE1 vs KE3
p1 <- ggplot() +
  # predicted data
  geom_point(data = response_df, aes(x = KE1, y = KE3, color = "drm"), size = 2) +
  geom_smooth(data = response_df, aes(x = KE1, y = p_KE3, color = "rrm", linetype = "rrm"), method = "gam", formula = y ~ s(x, k = 50), se = FALSE, color = "blue") +
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score1, y = eg_score3, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_277, y = mean_eg_hRPTECTERT1_50, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_277,
                    ymin = mean_eg_hRPTECTERT1_50 - SD_eg_hRPTECTERT1_50,
                    ymax = mean_eg_hRPTECTERT1_50 + SD_eg_hRPTECTERT1_50,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_50,
                     xmin = mean_eg_hRPTECTERT1_277 - SD_eg_hRPTECTERT1_277,
                     xmax = mean_eg_hRPTECTERT1_277 + SD_eg_hRPTECTERT1_277,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Inhibition, mtETC Chain Complexes VS Increased, oxidative stress", x = "Inhibition, mtETC Chain Complexes", y = "Increased, oxidative stress", color = "DOSE") +
  # facet_zoom(y = p_KE3 < 25) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  theme_minimal()

# plot2: KE2 vs KE3
p2 <- ggplot() +
  # predicted data
   geom_point(data = response_df, aes(x = KE2, y = KE3, color = "drm"), size = 2) +
  geom_smooth(data = response_df, aes(x = KE2, y = p_KE3, color = "rrm", linetype = "rrm"), method = "gam", se = FALSE, color = "blue") + 
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score2, y = eg_score3, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_239, y = mean_eg_hRPTECTERT1_50, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_239,
                    ymin = mean_eg_hRPTECTERT1_50 - SD_eg_hRPTECTERT1_50,
                    ymax = mean_eg_hRPTECTERT1_50 + SD_eg_hRPTECTERT1_50,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_50,
                     xmin = mean_eg_hRPTECTERT1_239 - SD_eg_hRPTECTERT1_239,
                     xmax = mean_eg_hRPTECTERT1_239 + SD_eg_hRPTECTERT1_239,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Binding, Thiol/seleno proteins VS Increased, oxidative stress", x = "Binding, Thiol/seleno proteins", y = "Increased, oxidative stress", color = "DOSE") +
  # facet_zoom(y = p_KE3 < 25) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  theme_minimal()
# print plots
print(p1)
print(p2)
```

### save
```{r}
ggsave(filename = "../data/R_R/277_239_50/p1.pdf",
       plot = p1,
       width = 8,
       height = 6,
       dpi = 300)
ggsave(filename = "../data/R_R/277_239_50/p2.pdf",
       plot = p2,
       width = 8,
       height = 6,
       dpi = 300)
```