---
title: "R-R: 164 + 97 ->71"
output: html_document
date: "2025-07-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyr)
library(ggplot2)
library(dplyr)
```


### module 264 + module 97 = module 71
# KE3: Increase, Apoptosis--71
# KE1: Increased, Lipid peroxidation--264
# KE2: Activation of ER stress--97
### use data predicted from dose-response models
```{r}
# define LL.3 manually
LL3_function <- function(dose, b, d, e) {
  d / (1 + exp(b * (log(dose) - log(e))))
}
MM2_function <- function(dose, d, e) {
  d/(1 + (e/dose))
}
MM3_function <- function(dose, c, d, e) {
  c + (d - c)/(1 + (e/dose))
}
# define the range of dose, for model fitting
# doses <- 10^seq(log10(100), log10(40000000), length.out = 500)
# change to smaller range when plotting!!!! otherwise it's too large, the plots won't be beautifal
doses <- 10^seq(log10(100), log10(5000), length.out = 50)

# ceate an empty data frame
response_df <- data.frame(dose = doses)

# stimulate KE1 data, using LL.3 model
# response_df$KE1 <- LL3_function(doses, b = -17.85, d = 2.39, e = 921.64)
response_df$KE1 <- MM3_function(doses, c = -0.58, d = 200.7, e = 89671.97) # remove outlier

# stimulate KE2 data, using LL.3 model
response_df$KE2 <- MM2_function(doses, d = 8.64, e = 2353.44) # remove outlier
# response_df$KE2 <- LL3_function(doses, b = -13.9, d = 1.7, e = 271.06)

# stimulate KE3 data, using MM.2 model
response_df$KE3 <- MM2_function(doses, d = 13.19, e = 2995.21) # remove outlier
# response_df$KE3 <- LL3_function(doses, b = -1.47, d = 5.23, e = 701.66) # remove outlier
# response_df$KE3 <- LL3_function(doses, b = -2.63, d = 2.51, e = 302.87)

rownames(response_df) <- response_df[, 1] # set the first column as row names(dose)
response_df <- response_df[, -1] # delete the first column

print(response_df)
```

# examine the data, check if the predictor variables have a linear association with the response variable
```{r}
pairs(response_df, pch = 18, col = "steelblue")
ggpairs(response_df)
```

# fit model
```{r}
model <- nls(KE3 ~ a * KE1 / (K + KE1) + b * KE2, data = response_df, start = list(a = 0.74, K = 2.5, b = 0.4))
summary(model)
```

#### The following two chunks were used to check how well the fitted model matches the data used for fitting.
#### when dealing with complex models that are hard to fit, these plots help in selecting initial parameter estimates that closely align with the data used for fitting.
```{r}
response_function <- function(KE1, KE2, a, k, b){
  a * KE1 / (k + KE1) + b * KE2
}
# set parameters
a <- 3.37
k <- 13.85
b <- 1.16

# predict KE3
response_df <- response_df %>%
  mutate(p_KE3 = response_function(KE1, KE2, a, k, b))

print(response_df)
```
### data type: data predicted from assumed equations VS data predicted from dose-response models
```{r}
# plot1: KE1 vs KE3
p1 <- ggplot() +
  # predicted data
  geom_point(data = response_df, aes(x = KE1, y = p_KE3, color = "Predicted", shape = "Predicted"), size = 2) +
  # geom_smooth(data = response_df, aes(x = KE1, y = p_KE3), method = "loess", se = FALSE, color = "blue") +
  geom_smooth(data = response_df, aes(x = KE1, y = p_KE3), method = "gam", se = FALSE, color = "blue") + # smoother method
  # experimental data
  geom_point(data = response_df, aes(x = KE1, y = KE3, color = "Predicted from drm", shape = "Predicted from drm"), size = 2) +
  scale_color_manual(name = "Data Type", values = c("Predicted" = "red", "Predicted from drm" = "red")) +
  scale_shape_manual(name = "Data Type", values = c("Predicted" = 16, "Predicted from drm" = 1)) +
  theme_minimal() +
  #facet_zoom(y = p_KE3 < 30) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  labs(title = "KE1 vs KE3", x = "KE1", y = "KE3")

# plot2: KE2 vs KE3
p2 <- ggplot(response_df, aes(x = KE2)) +
  # predicted data
   geom_point(data = response_df, aes(x = KE2, y = p_KE3, color = "Predicted", shape = "Predicted"), size = 2) +
  # geom_smooth(data = response_df, aes(x = KE2, y = p_KE3), method = "loess", se = FALSE, color = "blue") +
  geom_smooth(data = response_df, aes(x = KE2, y = p_KE3), method = "gam", se = FALSE, color = "blue") +
  # experimental data
  geom_point(data = response_df, aes(x = KE2, y = KE3, color = "Predicted from drm", shape = "Predicted from drm"), size = 2) +
  scale_color_manual(name = "Data Type", values = c("Predicted" = "red", "Predicted from drm" = "red")) +
  scale_shape_manual(name = "Data Type", values = c("Predicted" = 16, "Predicted from drm" = 1)) +
  theme_minimal() +
  # facet_zoom(y = p_KE3 < 30) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  labs(title = "KE2 vs KE3", x = "KE2", y = "KE3")

# print plots
print(p1)
print(p2)
```

### extract experimental data # for error bar in plots
```{r}
datafolder = "C:/Users/ZTian/Desktop/RP1/data/moduleTable.txt"
module_table <- read.delim(datafolder)
module_table <- module_table %>%
  mutate(conc_level = case_when(
    conc_level == 1 ~ 250,
    conc_level == 2 ~ 500,
    conc_level == 3 ~ 1000
  ))
KE1 <- module_table %>% dplyr::filter(
    module == 'hRPTECTERT1_264' & # Increased, Lipid peroxidation
      time == 24 &
      experiment %in% c(
        'uploadFile1_GENTAMYCINluhumanrptectert1',
        'uploadFile2_GENTAMYCINluhumanrptectert1',
        'uploadFile3_GENTAMYCINluhumanrptectert1'
      )
  )
  print(KE1)

  KE2 <- module_table %>% dplyr::filter(
    module == 'hRPTECTERT1_97' & # Activation of ER stress
      time == 24 &
      experiment %in% c(
        'uploadFile1_GENTAMYCINluhumanrptectert1',
        'uploadFile2_GENTAMYCINluhumanrptectert1',
        'uploadFile3_GENTAMYCINluhumanrptectert1'
      )
  )
  print(KE2)
  
  KE3 <- module_table %>% dplyr::filter(
    module == 'hRPTECTERT1_71' & # Increase, Apoptosis
      time == 24 &
      experiment %in% c(
        'uploadFile1_GENTAMYCINluhumanrptectert1',
        'uploadFile2_GENTAMYCINluhumanrptectert1',
        'uploadFile3_GENTAMYCINluhumanrptectert1'
      )
  )
  print(KE3)
  
  # simplify these data frame and merge them
  KE1_filtered <- KE1 %>%
    select(experiment, conc_level, eg_score) %>%
    rename(eg_score1 = eg_score)
  
  KE2_filtered <- KE2 %>%
    select(experiment, conc_level, eg_score) %>%
    rename(eg_score2 = eg_score)
  
  KE3_filtered <- KE3 %>%
    select(experiment, conc_level, eg_score) %>%
    rename(eg_score3 = eg_score)
  
  merged_data <- KE1_filtered %>%
  full_join(KE2_filtered, by = c("experiment", "conc_level")) %>%
  full_join(KE3_filtered, by = c("experiment", "conc_level"))
  print(merged_data)

```

### final plotting
```{r}
# KE3: Increase, Apoptosis--71
# KE1: Increased, Lipid peroxidation--264
# KE2: Activation of ER stress--97
# prepare the data frame used to draw error bars
# RStudio's table view (View()) only displays rounded numbers, not the actual stored values.
summarized_moduletable = module_table %>% filter(time == 24) %>%
  # # remove outlier from error bar
  filter(!(module == "hRPTECTERT1_97" & conc_level == 1000 &
           startsWith(as.character(eg_score), "-0.527544"))) %>%
  filter(!(module == "hRPTECTERT1_264" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.035441"))) %>%
  filter(!(module == "hRPTECTERT1_71" & conc_level == 1000 &
           startsWith(as.character(eg_score), "0.647139"))) %>%
  group_by(module, conc_level) %>%
  summarize(mean_eg = mean(eg_score, na.rm = TRUE), SD_eg = sd(eg_score, na.rm = TRUE)) %>%
  pivot_wider(names_from = module, values_from = c(mean_eg, SD_eg))
# # remove outlier from experimental data points
#   merged_data <- merged_data[-7, ]
#   # filter(!(eg_score2 == -0.5275449 & eg_score3 == 0.6471399)), cannot delete the row due to floating-point precision

# plot1: KE1 vs KE3
p1 <- ggplot() +
  # # predicted data
  geom_point(data = response_df, aes(x = KE1, y = KE3, color = "drm"), size = 2) +
  geom_smooth(data = response_df, aes(x = KE1, y = p_KE3, color = "rrm", linetype = "rrm"), method = "gam", se = FALSE, color = "blue") +
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score1, y = eg_score3, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_264, y = mean_eg_hRPTECTERT1_71, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_264,
                    ymin = mean_eg_hRPTECTERT1_71 - SD_eg_hRPTECTERT1_71,
                    ymax = mean_eg_hRPTECTERT1_71 + SD_eg_hRPTECTERT1_71,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_71,
                     xmin = mean_eg_hRPTECTERT1_264 - SD_eg_hRPTECTERT1_264,
                     xmax = mean_eg_hRPTECTERT1_264 + SD_eg_hRPTECTERT1_264,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Increased, Lipid peroxidation VS Increase, Apoptosis", x = "Increased, Lipid peroxidation", y = "Increase, Apoptosis", color = "DOSE") +
  # facet_zoom(y = p_KE3 < 25) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  theme_minimal()

# plot2: KE2 vs KE3
p2 <- ggplot() +
  # # predicted data
  geom_point(data = response_df, aes(x = KE2, y = KE3, color = "drm"), size = 2) +
  geom_smooth(data = response_df, aes(x = KE2, y = p_KE3, color = "rrm", linetype = "rrm"), method = "gam", se = FALSE, color = "blue") +
  # experimental data points
  geom_point(data = merged_data, aes(x = eg_score2, y = eg_score3, color = as.factor(conc_level)), alpha = 0.7) +
  # error bar
  geom_point(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_97, y = mean_eg_hRPTECTERT1_71, color = as.factor(conc_level)), size = 2) +
  geom_errorbar(data = summarized_moduletable, aes(x = mean_eg_hRPTECTERT1_97,
                    ymin = mean_eg_hRPTECTERT1_71 - SD_eg_hRPTECTERT1_71,
                    ymax = mean_eg_hRPTECTERT1_71 + SD_eg_hRPTECTERT1_71,
                    color = as.factor(conc_level)), width = 0.1) +
  geom_errorbarh(data = summarized_moduletable, aes(y = mean_eg_hRPTECTERT1_71,
                     xmin = mean_eg_hRPTECTERT1_97 - SD_eg_hRPTECTERT1_97,
                     xmax = mean_eg_hRPTECTERT1_97 + SD_eg_hRPTECTERT1_97,
                     color = as.factor(conc_level)), height = 0.1) +
  
  labs(title = "Activation of ER stress VS Increase, Apoptosis", x = "Activation of ER stress", y = "Increase, Apoptosis", color = "DOSE") +
  # facet_zoom(y = p_KE3 < 25) + # if there are extreme data, plot will be strange. This is to exclude extreme data from plot
  theme_minimal()
# print plots
print(p1)
print(p2)
```
### save
```{r}
ggsave(filename = "../data/R_R/164_97_71/p1.pdf",
       plot = p1,
       width = 8,
       height = 6,
       dpi = 300)
ggsave(filename = "../data/R_R/164_97_71/p2.pdf",
       plot = p2,
       width = 8,
       height = 6,
       dpi = 300)
```

